---
title: "rds analysis - single df: draft 1"
output: html_notebook
---


```{r}
# load in the data (1 data frame)
rds_dir <- "C:/Users/Jabba/OneDrive - University of Ottawa/MSc/Projects/Varying Normal Cells/rds_files/"

rds_file_paths <- list.files(path = rds_dir, full.names = TRUE)

example_rds <- readRDS(rds_file_paths[20])
```

# RDS Analysis
## TH cells
```{r}
condition <- tools::file_path_sans_ext(basename(rds_file_paths[20]))

th_rows <- grepl("TH1|TH2", rownames(example_rds))
total_th_cells <- sum(th_rows)
```

### TH NA cells
```{r}
th_NA_cells <- sum(th_rows & is.na(example_rds$class))

# calculate % of TH cells that have class = NA
percent_Allth_NA <- (th_NA_cells / total_th_cells)*100
```


### TH filtered
```{r}
th_filtered_cells <- sum(th_rows & example_rds$class == "filtered" & !is.na(example_rds$class))

# calculate % of TH cells that have class = filtered
percent_Allth_filtered <- (th_filtered_cells / total_th_cells)*100
```

### Determine number & percent of TH cells that get classified
```{r}
# number of TH cells classified:
th_classified_rows <- (grepl("TH1|TH2", rownames(example_rds)) & !example_rds$class=="filtered" & !is.na(example_rds$class))

num_th_classified <- sum(th_classified_rows)

# percent of TH cells classified, out of all of the TH cells
percent_th_classified <- num_th_classified/total_th_cells*100
```


### TH tumor cells
```{r}
th_tumor_cells <- sum(th_rows & example_rds$class == "tumor" & !is.na(example_rds$class))
# ^ checked by visually counting and is correct


percent_th_tumor <- (th_tumor_cells / num_th_classified) * 100
# good
```


### TH normal cells
```{r}
th_normal_cells <- sum(th_rows & example_rds$class == "normal" & !is.na(example_rds$class))

percent_th_normal <- (th_normal_cells / num_th_classified) * 100
```


## Leuk Cells. NOTE: when iterating, for PB2: can just swap sample name. For combined: can just add what was written in PB2.
```{r}
leuk_rows <- grepl("BM4|PB2", rownames(example_rds))
total_leuk_cells <- sum(leuk_rows)
```


### leuk NA
```{r}
leuk_NA_cells <- sum(leuk_rows & is.na(example_rds$class))

percent_leuk_NA <- (leuk_NA_cells / total_leuk_cells)*100
```

### leuk filtered
```{r}
leuk_filtered_cells <- sum(leuk_rows & example_rds$class == "filtered" & !is.na(example_rds$class))

percent_leuk_filtered <- (leuk_filtered_cells / total_leuk_cells)*100
```

### Determine number & percent of leuk cells that get classified
```{r}
# number of leuk cells classified:
leuk_classified_rows <- (grepl("BM4|PB2", rownames(example_rds)) & !example_rds$class=="filtered" & !is.na(example_rds$class))

num_leuk_classified <- sum(leuk_classified_rows)

# percent of leuk cells classified, out of all of the leuk cells
percent_leuk_classified <- num_leuk_classified/total_leuk_cells*100
```


### leuk tumor
```{r}
leuk_tumor_cells <- sum(leuk_rows & example_rds$class == "tumor" & !is.na(example_rds$class))

percent_leuk_tumor <- leuk_tumor_cells/num_leuk_classified*100
```

### leuk normal
```{r}
leuk_normal_cells <- sum(leuk_rows & example_rds$class == "normal" & !is.na(example_rds$class))

percent_leuk_normal <- leuk_normal_cells/num_leuk_classified*100
```

## prepare data frame containing the percentage results
```{r}
results_df_perc <- data.frame(
    Condition = condition,
    All_TH_Percent_NA = percent_Allth_NA,
    All_TH_Percent_Filtered = percent_Allth_filtered,
    classified_TH_Percent_Normal = percent_th_normal,
    Classified_TH_Percent_Tumor = percent_th_tumor,
    All_leuk_Percent_NA = percent_leuk_NA,
    All_leuk_Percent_Filtered = percent_leuk_filtered,
    Classified_leuk_Percent_Tumor = percent_leuk_tumor,
    Classified_leuk_Percent_Normal = percent_leuk_normal
  )
```

## prepare data frame containing the count results
```{r}
th1_rows <- grepl("TH1", rownames(example_rds))
total_th1_cells <- sum(th1_rows)

th2_rows <- grepl("TH2", rownames(example_rds))
total_th2_cells <- sum(th2_rows)

#Question: would it make sense to separate counts and percentages into different data frames? Or would it be better to just have everything in one?
results_df_num <- data.frame(
  Condition = condition,
  Total_TH1 = total_th1_cells,
  Total_TH2 = total_th2_cells,
  Total_TH = total_th_cells,
  Number_TH_NA = th_NA_cells, 
  Number_TH_Filtered = th_filtered_cells,
  Number_TH_classified = num_th_classified,
  Number_TH_Tumor = th_tumor_cells,
  Number_TH_Normal = th_normal_cells,
  Total_Leuk = total_leuk_cells,
  Number_Leuk_NA = leuk_NA_cells,
  Number_Leuk_Filtered = leuk_filtered_cells,
  Number_Leuk_Classified = num_leuk_classified,
  Number_Leuk_Tumor = leuk_tumor_cells,
  Number_Leuk_Normal = leuk_normal_cells
)
```

