---
title: "rds analysis - draft 3"
output: html_notebook
---

# initialize
```{r}
# load in the data
rds_dir <- "C:/Users/Jabba/OneDrive - University of Ottawa/MSc/Projects/Varying Normal Cells/rds_files/"

rds_file_paths <- list.files(path = rds_dir, full.names = TRUE)
```

```{r}
# Read all .rds files into a list
data_list <- lapply(rds_file_paths, readRDS)

# name the list elements 
names(data_list) <- tools::file_path_sans_ext(basename(rds_file_paths))
```

# Create the function that will do the analysis to each  
```{r}
classification_analysis <- function(x) {
# start the analysis
  ## TH cells

  ### Total TH cells
  th_rows <- grepl("TH1|TH2", rownames(x))
  total_th_cells <- sum(th_rows)
  
  #### TH NA cells
  th_NA_cells <- sum(th_rows & is.na(x$class))
  
  ##### calculate % of TH cells that have class = NA
  percent_Allth_NA <- (th_NA_cells / total_th_cells)*100
  
  #### TH filtered cells
  th_filtered_cells <- sum(th_rows & x$class == "filtered" & !is.na(x$class))
  
  percent_Allth_filtered <- (th_filtered_cells / total_th_cells)*100
  
  #### number of TH cells classified:
  th_classified_rows <- (grepl("TH1|TH2", rownames(x)) & !x$class=="filtered" & !is.na(x$class))

  num_th_classified <- sum(th_classified_rows)

  #### percent of TH cells classified, out of all of the TH cells
  percent_th_classified <- num_th_classified/total_th_cells*100
  
  #### TH tumor cells
  th_tumor_cells <- sum(th_rows & x$class == "tumor" & !is.na(x$class))
  
  percent_classified_th_tumor <- th_tumor_cells/num_th_classified*100
  
  #### TH normal cells
    th_normal_cells <- sum(th_rows & x$class == "normal" & !is.na(x$class))
  
  percent_classified_th_normal <- th_normal_cells/num_th_classified*100
  
  
  
  
  ### TH1 cells
  th1_rows <- grepl("TH1", rownames(x))
  total_th1_cells <- sum(th1_rows)
  
  #### th1 NA cells
  th1_NA_cells <- sum(th1_rows & is.na(x$class))
  ##### calculate % of th1 cells that have class = NA
  percent_Allth1_NA <- (th1_NA_cells / total_th1_cells)*100
  
  #### th1 filtered cells
  th1_filtered_cells <- sum(th1_rows & x$class == "filtered" & !is.na(x$class))
  
  percent_Allth1_filtered <- (th1_filtered_cells / total_th1_cells)*100
  
  #### number of th1 cells classified:
  th1_classified_rows <- (grepl("TH1", rownames(x)) & !x$class=="filtered" & !is.na(x$class))

  num_th1_classified <- sum(th1_classified_rows)

  #### percent of th1 cells classified, out of all of the th1 cells
  percent_th1_classified <- num_th1_classified/total_th1_cells*100
  
  #### th1 tumor cells
  th1_tumor_cells <- sum(th1_rows & x$class == "tumor" & !is.na(x$class))
  
  percent_classified_th1_tumor <- th1_tumor_cells/num_th1_classified*100
  
  #### th1 normal cells
    th1_normal_cells <- sum(th1_rows & x$class == "normal" & !is.na(x$class))
  
  percent_classified_th1_normal <- th1_normal_cells/num_th1_classified*100
  
  
  
  ### th2 cells
  th2_rows <- grepl("TH2", rownames(x))
  total_th2_cells <- sum(th2_rows)
  
  #### th2 NA cells
  th2_NA_cells <- sum(th2_rows & is.na(x$class))
  ##### calculate % of th2 cells that have class = NA
  percent_Allth2_NA <- (th2_NA_cells / total_th2_cells)*100
  
  #### th2 filtered cells
  th2_filtered_cells <- sum(th2_rows & x$class == "filtered" & !is.na(x$class))
  
  percent_Allth2_filtered <- (th2_filtered_cells / total_th2_cells)*100
  
  #### number of th2 cells classified:
  th2_classified_rows <- (grepl("TH2", rownames(x)) & !x$class=="filtered" & !is.na(x$class))

  num_th2_classified <- sum(th2_classified_rows)

  #### percent of th2 cells classified, out of all of the th2 cells
  percent_th2_classified <- num_th2_classified/total_th2_cells*100
  
  #### th2 tumor cells
  th2_tumor_cells <- sum(th2_rows & x$class == "tumor" & !is.na(x$class))
  
  percent_classified_th2_tumor <- th2_tumor_cells/num_th2_classified*100
  
  #### th2 normal cells
    th2_normal_cells <- sum(th2_rows & x$class == "normal" & !is.na(x$class))
  
  percent_classified_th2_normal <- th2_normal_cells/num_th2_classified*100
  
  
  
  ### Total Leuk cells
if (any(grepl("BM4", rownames(x))) && !any(grepl("PB2", rownames(x)))) {
  total_leuk_cells <- 7493
} else if (any(grepl("PB2", rownames(x))) && !any(grepl("BM4", rownames(x)))) {
  total_leuk_cells <- 7150
} else {
  total_leuk_cells <- 14643
}
  
  leuk_rows <- grepl("BM4|PB2", rownames(x))
  
  #### leuk NA cells
  leuk_NA_cells <- sum(leuk_rows & is.na(x$class))
  ##### calculate % of leuk cells that have class = NA
  percent_Allleuk_NA <- (leuk_NA_cells / total_leuk_cells)*100
  
  #### leuk filtered cells
  leuk_filtered_cells <- sum(leuk_rows & x$class == "filtered" & !is.na(x$class))
  
  percent_Allleuk_filtered <- (leuk_filtered_cells / total_leuk_cells)*100
  
  #### number of leuk cells classified:
  leuk_classified_rows <- (grepl("BM4|PB2", rownames(x)) & !x$class=="filtered" & !is.na(x$class))

  num_leuk_classified <- sum(leuk_classified_rows)

  #### percent of leuk cells classified, out of all of the leuk cells
  percent_leuk_classified <- num_leuk_classified/total_leuk_cells*100
  
  #### leuk tumor cells
  leuk_tumor_cells <- sum(leuk_rows & x$class == "tumor" & !is.na(x$class))
  
  percent_classified_leuk_tumor <- leuk_tumor_cells/num_leuk_classified*100
  
  #### leuk normal cells
    leuk_normal_cells <- sum(leuk_rows & x$class == "normal" & !is.na(x$class))
  
  percent_classified_leuk_normal <- leuk_normal_cells/num_leuk_classified*100

if (any(grepl("BM4", rownames(x))) && !any(grepl("PB2", rownames(x)))) {
  
  ### BM4 cells
  bm4_rows <- grepl("BM4", rownames(x))
  total_bm4_cells <- 7493
  
  #### bm4 NA cells
  bm4_NA_cells <- sum(bm4_rows & is.na(x$class))
  ##### calculate % of bm4 cells that have class = NA
  percent_Allbm4_NA <- (bm4_NA_cells / total_bm4_cells)*100
  
  #### bm4 filtered cells
  bm4_filtered_cells <- sum(bm4_rows & x$class == "filtered" & !is.na(x$class))
  
  percent_Allbm4_filtered <- (bm4_filtered_cells / total_bm4_cells)*100
  
  #### number of bm4 cells classified:
  bm4_classified_rows <- (grepl("BM4", rownames(x)) & !x$class=="filtered" & !is.na(x$class))

  num_bm4_classified <- sum(bm4_classified_rows)

  #### percent of bm4 cells classified, out of all of the bm4 cells
  percent_bm4_classified <- num_bm4_classified/total_bm4_cells*100
  
  #### bm4 tumor cells
  bm4_tumor_cells <- sum(bm4_rows & x$class == "tumor" & !is.na(x$class))
  
  percent_classified_bm4_tumor <- bm4_tumor_cells/num_bm4_classified*100
  
  #### bm4 normal cells
    bm4_normal_cells <- sum(bm4_rows & x$class == "normal" & !is.na(x$class))
  
  percent_classified_bm4_normal <- bm4_normal_cells/num_bm4_classified*100
 
percent_Allpb2_NA <- 0
percent_Allpb2_filtered <- 0
percent_pb2_classified <- 0
percent_classified_pb2_tumor <- 0
percent_classified_pb2_normal <- 0

selected_analysis_bm4 <- "Executed BM4-specific analysis"  
  
} else if (any(grepl("PB2", rownames(x))) && !any(grepl("BM4", rownames(x)))) {
  
  ### pb2 cells
  pb2_rows <- grepl("PB2", rownames(x))
  total_pb2_cells <- 7150
  
  #### pb2 NA cells
  pb2_NA_cells <- sum(pb2_rows & is.na(x$class))
  ##### calculate % of pb2 cells that have class = NA
  percent_Allpb2_NA <- (pb2_NA_cells / total_pb2_cells)*100
  
  #### pb2 filtered cells
  pb2_filtered_cells <- sum(pb2_rows & x$class == "filtered" & !is.na(x$class))
  
  percent_Allpb2_filtered <- (pb2_filtered_cells / total_pb2_cells)*100
  
  #### number of pb2 cells classified:
  pb2_classified_rows <- (grepl("PB2", rownames(x)) & !x$class=="filtered" & !is.na(x$class))

  num_pb2_classified <- sum(pb2_classified_rows)

  #### percent of pb2 cells classified, out of all of the pb2 cells
  percent_pb2_classified <- num_pb2_classified/total_pb2_cells*100
  
  #### pb2 tumor cells
  pb2_tumor_cells <- sum(pb2_rows & x$class == "tumor" & !is.na(x$class))
  
  percent_classified_pb2_tumor <- pb2_tumor_cells/num_pb2_classified*100
  
  #### pb2 normal cells
    pb2_normal_cells <- sum(pb2_rows & x$class == "normal" & !is.na(x$class))
  
  percent_classified_pb2_normal <- pb2_normal_cells/num_pb2_classified*100
  
  
percent_Allbm4_NA <- 0
percent_Allbm4_filtered <- 0
percent_bm4_classified <- 0
percent_classified_bm4_tumor <- 0
percent_classified_bm4_normal <- 0

selected_analysis_pb2 <- "Executed PB2-specific analysis"
  
} else {
  ### BM4 cells
  bm4_rows <- grepl("BM4", rownames(x))
  total_bm4_cells <- 7493
  
  #### bm4 NA cells
  bm4_NA_cells <- sum(bm4_rows & is.na(x$class))
  ##### calculate % of bm4 cells that have class = NA
  percent_Allbm4_NA <- (bm4_NA_cells / total_bm4_cells)*100
  
  #### bm4 filtered cells
  bm4_filtered_cells <- sum(bm4_rows & x$class == "filtered" & !is.na(x$class))
  
  percent_Allbm4_filtered <- (bm4_filtered_cells / total_bm4_cells)*100
  
  #### number of bm4 cells classified:
  bm4_classified_rows <- (grepl("BM4", rownames(x)) & !x$class=="filtered" & !is.na(x$class))

  num_bm4_classified <- sum(bm4_classified_rows)

  #### percent of bm4 cells classified, out of all of the bm4 cells
  percent_bm4_classified <- num_bm4_classified/total_bm4_cells*100
  
  #### bm4 tumor cells
  bm4_tumor_cells <- sum(bm4_rows & x$class == "tumor" & !is.na(x$class))
  
  percent_classified_bm4_tumor <- bm4_tumor_cells/num_bm4_classified*100
  
  #### bm4 normal cells
    bm4_normal_cells <- sum(bm4_rows & x$class == "normal" & !is.na(x$class))
  
  percent_classified_bm4_normal <- bm4_normal_cells/num_bm4_classified*100
  
  
  
  ### pb2 cells
  pb2_rows <- grepl("PB2", rownames(x))
  total_pb2_cells <- 7150
  
  #### pb2 NA cells
  pb2_NA_cells <- sum(pb2_rows & is.na(x$class))
  ##### calculate % of pb2 cells that have class = NA
  percent_Allpb2_NA <- (pb2_NA_cells / total_pb2_cells)*100
  
  #### pb2 filtered cells
  pb2_filtered_cells <- sum(pb2_rows & x$class == "filtered" & !is.na(x$class))
  
  percent_Allpb2_filtered <- (pb2_filtered_cells / total_pb2_cells)*100
  
  #### number of pb2 cells classified:
  pb2_classified_rows <- (grepl("PB2", rownames(x)) & !x$class=="filtered" & !is.na(x$class))

  num_pb2_classified <- sum(pb2_classified_rows)

  #### percent of pb2 cells classified, out of all of the pb2 cells
  percent_pb2_classified <- num_pb2_classified/total_pb2_cells*100
  
  #### pb2 tumor cells
  pb2_tumor_cells <- sum(pb2_rows & x$class == "tumor" & !is.na(x$class))
  
  percent_classified_pb2_tumor <- pb2_tumor_cells/num_pb2_classified*100
  
  #### pb2 normal cells
    pb2_normal_cells <- sum(pb2_rows & x$class == "normal" & !is.na(x$class))
  
  percent_classified_pb2_normal <- pb2_normal_cells/num_pb2_classified*100

  selected_analysis_combined <- "Executed combined analysis"
  }

results_df_perc <- data.frame(
    All_TH_Percent_NA = percent_Allth_NA,
    All_TH_Percent_Filtered = percent_Allth_filtered,
    All_TH_Percent_Classified = percent_th_classified,
    Classified_TH_Percent_Tumor = percent_classified_th_tumor,
    Classified_TH_Percent_Normal = percent_classified_th_normal,
    
    All_TH1_Percent_NA = percent_Allth1_NA,
    All_TH1_Percent_Filtered = percent_Allth1_filtered,
    All_TH1_Percent_Classified = percent_th1_classified,
    Classified_TH1_Percent_Tumor = percent_classified_th1_tumor,
    Classified_TH1_Percent_Normal = percent_classified_th1_normal,
    
    All_TH2_Percent_NA = percent_Allth2_NA,
    All_TH2_Percent_Filtered = percent_Allth2_filtered,
    All_TH2_Percent_Classified = percent_th2_classified,
    Classified_TH2_Percent_Tumor = percent_classified_th2_tumor,
    Classified_TH2_Percent_Normal = percent_classified_th2_normal,
    
    All_Leuk_Percent_NA = percent_Allleuk_NA,
    All_Leuk_Percent_Filtered = percent_Allleuk_filtered,
    All_Leuk_Percent_Classified = percent_leuk_classified,
    Classified_Leuk_Percent_Tumor = percent_classified_leuk_tumor,
    Classified_Leuk_Percent_Normal = percent_classified_leuk_normal,
    
    All_BM4_Percent_NA = percent_Allbm4_NA,
    All_BM4_Percent_Filtered = percent_Allbm4_filtered,
    All_BM4_Percent_Classified = percent_bm4_classified,
    Classified_BM4_Percent_Tumor = percent_classified_bm4_tumor,
    Classified_BM4_Percent_Normal = percent_classified_bm4_normal,
    
    All_PB2_Percent_NA = percent_Allpb2_NA,
    All_PB2_Percent_Filtered = percent_Allpb2_filtered,
    All_PB2_Percent_Classified = percent_pb2_classified,
    Classified_PB2_Percent_Tumor = percent_classified_pb2_tumor,
    classified_PB2_Percent_Normal = percent_classified_pb2_normal
  )
}
```

# use analysis fxn on full list (i.e. all datasets)
```{r}
results_list <- lapply(data_list, classification_analysis)

results_df <- do.call(rbind, results_list)
```

# order results_df
```{r}
install.packages('gtools')
library(gtools)

results_df <- results_df[mixedsort(rownames(results_df)), ]
```

# plots
## scatter plot of the percentage of classified thymus cells that get classified as normal
### install dependencies
```{r}
install.packages('tidyverse')
library(tidyverse)
```

### prep the inputs
```{r}
total_th1_count <- 5064
total_th2_count <- 3253
total_th_count <- total_th1_count+total_th2_count
```

#### create a fxn that determines the correct # cells given a string (condition) as input
```{r}
cond_names <- rownames(results_df)

## from copilot:

# Function to convert the strings to numerical values
convert_to_numeric <- function(cond_names, total_th_count) {
  sapply(cond_names, function(x) {
    # Extract the numeric part
    num <- as.numeric(sub(".*_", "", x))
    
    # Check if it's a count or percentage and convert accordingly
    if (grepl("count", x)) {
      return(num)
    } else if (grepl("perc", x)) {
      return((num / 100) * total_th_count)
    } else {
      return(NA)  # Handle unexpected cases
    }
  })
}
```

#### apply the fxn to calculate (roughly) the proper amt of normal cells in each rds file
```{r}
num_normal_cells <- convert_to_numeric(cond_names, total_th_count)

# Print the result
print(num_normal_cells)
```

#### define the y-axis variable (% of classified TH cells that get classified as tumor)
```{r}
classified_th_perc_normal <- results_df$Classified_TH_Percent_Normal

#check names
names(num_normal_cells) == rownames(results_df)

# define names for consistency
names(classified_th_perc_normal) <- names(num_normal_cells)


print(classified_th_perc_normal)
```

### create data frame for input to ggplot2
```{r}
ggplot2_input <- data.frame(num_normal_cells, classified_th_perc_normal)

ggplot2_input$Sample <- toupper(sub("_.*", "", rownames(ggplot2_input)))

desired_order <- c("BM4", "PB2", "COMBINED")

# Convert the prefix column to a factor with the specified levels
ggplot2_input$Sample <- factor(ggplot2_input$Sample, levels = desired_order)

# Reorder the data frame based on the Sample column
ggplot2_input <- ggplot2_input[order(ggplot2_input$Sample), ]
```


### plotting time!
```{r}
ggplot(ggplot2_input, aes(x = num_normal_cells, y = classified_th_perc_normal, color = Sample)) +
  geom_point() +
  labs(title = "Sample-specific effects observed on classification accuracy", x = "# normal thymus", y = "% normal cells classified correctly") + 
  theme_minimal()
```

### add lines of best fit
```{r}
ggplot(ggplot2_input, aes(x = num_normal_cells, y = classified_th_perc_normal, color = Sample)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +  # Add lines of best fit
  labs(title = "Sample-specific effects observed on classification accuracy", 
       x = "# normal thymus", 
       y = "% normal cells classified correctly") + 
  theme_minimal()
```

### Investigate the linear models
```{r}
library(dplyr)
library(broom) # For converting model outputs to data frames


# Fit linear models for each group and extract statistics
models <- ggplot2_input %>%
  group_by(Sample) %>%
  do(tidy(lm(classified_th_perc_normal ~ num_normal_cells, data = .))) %>%
  ungroup()

# View the coefficients and p-values
models

# Optionally, get additional statistics like R-squared
model_stats <- ggplot2_input %>%
  group_by(Sample) %>%
  do(glance(lm(classified_th_perc_normal ~ num_normal_cells, data = .))) %>%
  ungroup()

# View the model statistics, including R-squared
model_stats
```

```{r}

```
